Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-09-08 Manuscript/manuscript.tex
Small update to the section Stoich. binding in MS.

Binding stoichiometry of \ce{Ca2+} and POPC was 
thoroughly studied in the experimental work~\citenum{altenbach84}.
Several binding models were proposed and tested of which only one,
ternary complex binding model, provided a good fit of the experimental observations. 
In such a model, it is assumed that \ce{Ca2+} cations bind to a POPC membrane 
with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
Such a model also provides a good fit 
to our simulations with ECC-lipids 
(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
Simulations, however, allow us to directly evaluate the stoichiometry 
by calculating relative propensities of various \ce{Ca2+}:$n\times$POPC clusters
In Figure~\ref{fig:cacl_complexes} we see that ternary complex is indeed 
the most probable binding mode of calcium at 285~mM concentration.

Using our model, ECC-lipids, we fitted such a model to our simulations with a g
In such a model, there is a linear relationship between quantities 
$C_b$, mole fraction of bound \ce{Ca2+} per POPC, and $\sqrt(C_b/C_I)$, 
where $C_I$ is the concentration of free cations at the plane of ion binding~\cite{altenbach84}.
The concentration $C_b$ was obtained from an extrapolation of linear relation 
between deuterium NMR measurements and atomic absorption spectroscopy for low concetrations of \ce{CaCl2}.
Such an extrapolation is valid as long as the mode of \ce{Ca2+} binding 
remains constant throughout the extrapolation range. 
The concentration $C_I$ is determined by using the surface potential by using the Boltzmann equation.
However, Boltzmann theory yields inaccurate results
for divalent cations like \ce{Ca2+}~\cite{Andelmann1995}. 
An atomistic simulation, on the other hand, provides these quantities directly without severe assumptions. 
Hence we hypothesise that the discrepancy between the results in the experiment~\cite{altenbach84} and 
our simulations likely lays in the fact that 
the assumptions and relations used for determinig concentrations $C_b$ and $C_I$ in the experiment~\cite{altenbach84}
gradually do not hold for higher concentrations of \ce{Ca2+}.

Binding probability:
\ce{Ca2+} complex with
1 POPC - 13\%
2 POPC - 18\%
3 POPC - 12\%

@@

The good agreement of ECC-lipids model with experiments encourages us to analyse the
molecular binding details from MD simulations. Direct analysis of contacts between ions and
lipids from simulations suggest that the most abundant POPC:\ce{Ca^2+} complex 
has stoichiometry of 2~POPC:1~\ce{Ca^{2+}}.
As shown in Fig.~\ref{fig:cacl-bind} this is in agreement
with the ternary complex model suggested based on head group order parameter
experiments~\cite{altenbach84}.
In addition to the ternary complexes, there also is a non-negligable probability
of one \ce{Ca^{2+}} cross-bridging three POPC molecules.
Technical details of the analysis are in the SI.

>>>>

%Binding probability:
%\ce{Ca2+} complex with \\
%1 POPC - 13\% -- 30\% \\
%2 POPC - 18\% -- 42\% \\
%3 POPC - 12\% -- 28\% \\

Binding stoichiometry of \ce{Ca2+} and POPC was 
thoroughly studied in the experimental work~\citenum{altenbach84}.
Several binding models were proposed and tested of which only one,
ternary complex binding model, provided a good fit of the experimental observations. 
In such a model, it is assumed that \ce{Ca2+} cations bind to a POPC membrane 
with a stoichiometry 2~POPC:1~\ce{Ca^{2+}}.
Such a model also provides a good fit 
to our simulations with ECC-lipids 
(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
Simulations, however, allow us to directly evaluate the stoichiometry 
by calculating relative propensities of various \ce{Ca2+}:$n\times$POPC clusters
In Figure~\ref{fig:cacl_complexes} we see that ternary complex is indeed 
the most probable binding mode of calcium at 285~mM concentration. 
Apart from this complex, we also find complexes with 1~and 3~lipids
occuring with only a slightly lower but similar probability.
The fractions of \ce{Ca2+}:$n\times$POPC complexes at 285~mM concetration 
are then in order:
$42\%$ for two lipids,
$30\%$ for one lipid, and
$28\%$ for three lipids.
The symmetry of such a distribution --  
i.e. almost equal probabilities of complexes with 1 or 3 lipids 
that behave in the total average picture as complexes with 2 lipids -- 
provides clues why ternary complex binding model 
fits both simulation and experimental results relatively well, 
although it is apparently incorrect.

@@
